---
title: "Assignment 1 - Language Development in ASD - part 2"
author: '[R?ka Kereszt?nyi]'
date: '[9/12/2018]'
output:
  word_document: default
  html_document: default
---
    
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# you could optionally set include = TRUE globally while working on the exercise and then just change it to include = FALSE before you hand in
# knitr::opts_chunk$set(include = FALSE)
```

# Template for the hand-in
### Structure of the code chunks

Basic stuff:
- Loading the libraries
- Setting the directory and loading the data
- Look at the data (which variables are there? Are they in the right format?) and describe the participants (by diagnosis)

We will try to answer three questions:
    
- Do children with ASD develop language differently from non-ASD children?
- Do parents speak differently to children with ASD than to non-ASD ones?
- Which variables should we use to best explain the child linguistic performance?

### Loading the relevant libraries

Load necessary libraries : what will you need?

- e.g. something to plot with
- e.g. mixed effects models

```{r Load Libraries, include = FALSE}
#tidyverse, pacman, data.table, stringr, ggplot2, metrics, lmerTest, caret (rec)

library(ggplot2)  # data visualization
library(scales)   # date/time scales for plots
library(dplyr)    # data wrangling
library(tidyr)    # reshaping data
library(stringr)  # tools for strings
library(pastecs)  # descriptive statistics
library(lme4)     # mixed effect model

library(tidyverse)
library(psych)
library(RColorBrewer)
library(scales)
library(pacman)
library(lmerTest)
library(DHARMa)
```

### Define your working directory and load the data
If you created a project for this class and opened this Rmd file from within that project, your working directory is your project directory.

If you opened this Rmd file outside of a project, you will need some code to find the data:
- Create a new variable called locpath (localpath)
- Set it to be equal to your working directory
- Move to that directory (setwd(locpath))
- Load the data you saved last time (use read_csv(fileName))

If your're in a project, just put the data in the project folder and you're good to go! (Or make a data subfolder to keep it tidy around here)
```{r Load Data, include = FALSE}
setwd("C:/Users/Reka/Documents/EM3")
Data <- read.csv("data_exp2.csv") 
```

### Characterize the participants (Exercise 1)

Identify relevant variables: participants demographic characteristics, diagnosis, ADOS, Verbal IQ, Non Verbal IQ, Visit, Number of words used, Number of unique words used, length of utterance in both child and parents.

Make sure the variables are in the right format.

Describe the characteristics of the two groups of participants and whether the two groups are well matched.

```{r,background}
#exploring variables
#error:
Data$Ethnicity[Data$Ethnicity == "Bangledeshi"] <- "Bangladeshi"
#Summary - general
by(Data, Data$Diagnosis, summary)

#subset of Data
X <- Data %>% group_by(SUBJ, Ethnicity, Gender, Diagnosis, ADOS1, nonVerbalIQ1, verbalIQ1) %>% summarise(count = n())

#1.demographic background:
#A.diagnosis+ethnicity
summary(X$Ethnicity)
#graph
ggplot(X) + 
  aes(x = Ethnicity, stat="count") + 
  geom_bar() + 
  aes(x = Ethnicity, fill = Diagnosis) 

#B.diagnosis+gender
#summary(X$Diagnosis)
#summary(X$Gender)
summary(glm(Gender~Diagnosis, family = binomial, Data)) #gender balanced
#graph
ggplot(X) + 
  aes(x = Diagnosis, stat="count") + 
  geom_bar() + 
  aes(x = Diagnosis, fill = Gender) 

#C.diagnosis+age/visit
#summary(Data$Age) 
summary(lm(Age~Diagnosis, Data))
#graph
ggplot(Data) +
  geom_point(mapping = aes(x = VISIT, y = Age, color = Diagnosis)) + 
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
#age
TD <- subset(Data, Diagnosis=="TD")
ASD <- subset(Data, Diagnosis=="ASD")
t.test(TD$Age,ASD$Age) #difference in means is significant - Children with ASD seem to be older than healthy children.(reasonable)

```

```{r,ADOS+IQ}
#2.ADOS+diagnosis
#summary(X$ADOS1) 
#graph
ggplot(X) +
  aes(ADOS1, stat="count") + 
  geom_boxplot() + 
  aes(x = Diagnosis,y=ADOS1, fill = Diagnosis)
#compare 2 groups - difference in means is significant
#t.test(TD$ADOS1 , ASD$ADOS1)

#3.VISIT
##. 2  7  8  9 17 26 38 40 44 45 48 55 56  - no complete trials (13)

#4. Verbal + Non-verbal IQ
summary(lm(verbalIQ1~Diagnosis, X))
summary(lm(nonVerbalIQ1~Diagnosis, X))
#graph
#VERBALIQ
ggplot(X) +
  aes(verbalIQ1, stat="count") + 
  geom_boxplot() + 
  aes(Diagnosis,verbalIQ1, fill = Diagnosis)
#NONVERBALIQ
ggplot(X) +
  aes(nonVerbalIQ1, stat="count") + 
  geom_boxplot() + 
  aes(Diagnosis,nonVerbalIQ1, fill = Diagnosis)
#compare
TD1 <- subset(X, Diagnosis=="TD")
ASD1 <- subset(X, Diagnosis=="ASD")

t.test(TD1$verbalIQ1,ASD1$verbalIQ1) #better predictor than non-verbal IQ, but not significant (p>.05)
t.test(TD1$nonVerbalIQ1,ASD1$nonVerbalIQ1) #not significant (p>.05)
#gender/IQ graph
ggplot(X, aes(x = Diagnosis, y = verbalIQ1, fill = Diagnosis)) + geom_boxplot() + facet_wrap(~Gender)
ggplot(X, aes(x = Diagnosis, y = nonVerbalIQ1, fill = Diagnosis)) + geom_boxplot() + facet_wrap(~Gender)
```

```{r, types, token, MLU - mother/child}
#5. Number of words - tokens
#summary(Data$tokens_CHI) 
#compare: amount of words/TOKENS of mothers of both groups - no difference.
t.test(TD$tokens_MOT,ASD$tokens_MOT)
#compare: amount of words/TOKENS of CHILDREN of both groups
t.test(TD$tokens_CHI,ASD$tokens_CHI) #significant difference - autistic children seem to speak half ot the amount of words than 'normal' kids.

#plot - combined
ggplot(Data)+
  geom_boxplot(aes(x = Diagnosis, y = tokens_CHI, fill = Diagnosis)) +
  geom_boxplot(aes(x = Diagnosis, y = tokens_MOT)) +
  theme(legend.position="none")

#6. Unique words - types
#summary(Data$types_CHI) 
t.test(TD$types_MOT,ASD$types_MOT) #significant - between parents for both groups
t.test(TD$types_CHI,ASD$types_CHI) #significant - between children for noth groups
#plot - combined
ggplot(Data)+
  geom_boxplot(aes(x = Diagnosis, y = types_CHI, fill = Diagnosis)) +
  geom_boxplot(aes(x = Diagnosis, y = types_MOT)) +
  theme(legend.position="none")

#7. MLU - parents/children
t.test(TD$MOT_MLU,ASD$MOT_MLU) #significant - between parents for both groups
t.test(TD$CHI_MLU,ASD$CHI_MLU) #significant - between children for both groups
#plot - combined
ggplot(Data)+
  geom_boxplot(aes(x = Diagnosis, y = CHI_MLU, fill = Diagnosis)) +
  geom_boxplot(aes(x = Diagnosis, y = MOT_MLU)) +
  theme(legend.position="none")

```

```{r, with visit}
#LATERON
#graph - CHI tokens - over time
ggplot(Data, aes(x = VISIT, y = tokens_CHI, color = Diagnosis)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
#graph - CHI types - over time
ggplot(Data, aes(x = VISIT, y = types_CHI, color = Diagnosis)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
#graph - CHI MLU - over time
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = Diagnosis)) +
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
```

[REPORT THE RESULTS]

## Let's test hypothesis 1: Children with ASD display a language impairment  (Exercise 2)

### Hypothesis: The child's MLU changes: i) over time, ii) according to diagnosis

Let's start with a simple mixed effects linear model

Remember to plot the data first and then to run a statistical test.
- Which variable(s) should be included as fixed factors? - visit, Diagnosis
- Which variable(s) should be included as random factors? - kid/SUBJ, Ethnicity, Gender, 

```{r, model+ANOVA for comparison}
null_model = lmer(CHI_MLU~1+(1+VISIT|SUBJ), Data, REML=FALSE)
model2 = lmer(CHI_MLU~1+VISIT+Diagnosis + (1+VISIT|SUBJ), Data, REML=FALSE) 
#summary(model2)
model3 = lmer(CHI_MLU~1+VISIT*Diagnosis + (1+VISIT|SUBJ), Data, REML=FALSE)
#summary(model3)
anova(null_model, model2, model3 ) #model3 - with the interaction seems better?

#graph
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth() +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
```

Not too good, right? Let's check whether a growth curve model is better.
Remember: a growth curve model assesses whether changes in time can be described by linear, or quadratic, or cubic (or... etc.) components.
First build the different models, then compare them to see which one is better

```{r, model testing}
#growth curve models (polynomial)
quadratic = lmer(CHI_MLU ~ 1 + poly(VISIT,2) * Diagnosis + (1 + VISIT|SUBJ), Data) #degree2 
cubic = lmer(CHI_MLU ~ 1 + poly(VISIT,3) * Diagnosis + (1 + VISIT|SUBJ), Data) #degree3
quartic = lmer(CHI_MLU ~ 1 + poly(VISIT,4) * Diagnosis + (1 + VISIT|SUBJ), Data) #degree4

#summary(quadratic)
#summary(cubic)

anova(model3, quadratic, cubic, quartic) #(p<.001; p<.006, p>.05) cubic - AIC is the smallest, where p is still significant
```

```{r graphs, include = FALSE}
#linear
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth(method = lm) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")

#quadratic
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth(method = lm, formula = y ~ poly(x, 2)) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")

#cubic 
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth(method = lm, formula = y ~ poly(x, 3)) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
```

```{r, better model}
#cubic - better model
ggplot(Data, aes(x = VISIT, y = CHI_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth(method = lm, formula = y ~ poly(x, 3)) +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")

#cubic = lmer(CHI_MLU ~ 1 + poly(VISIT,3) * Diagnosis + (1 + VISIT|SUBJ), Data) #degree3
#summary(cubic)
```

Exciting right? 
Now it's time to report our results.
Remember to report:
- the estimates for each predictor (beta estimate, standard error, p-value)
- A plain word description of the results
- A plot of your best model's predictions

#RESULTS/hypothesis 1:
There is a significant effect of diagnosis, b= 0.65, SE = 0.65, t(61) = 4.07, p < .001, and time (VISIT), b=0.23,SE = 0.23, t(292) = 14.53, p < .0001, on the children's MLU (mean length of utterances).

## Let's test hypothesis 2: Parents speak equally to children with ASD and TD  (Exercise 3)

### Hypothesis: Parental MLU changes: i) over time, ii) according to diagnosis

```{r ex3, include = FALSE}
#model test
null_mot = lmer(MOT_MLU~1+(1+VISIT|SUBJ), Data)
mot2 = lmer(MOT_MLU~1+VISIT + Diagnosis + (1+VISIT|SUBJ), Data) #difference between visit (increase),diagnosis + mother MLU between groups. (TD higher a bit /0.5)
summary(mot2)
#mot3 = lmer(MOT_MLU~1+VISIT*Diagnosis + (1+VISIT|SUBJ), Data) #no interaction
#summary(mot3)
#anova(null_mot, mot2, mot3)

#graph
ggplot(Data, aes(x = VISIT, y = MOT_MLU, color = VISIT)) +
  geom_point() + 
  geom_smooth() +
  facet_wrap(~ Diagnosis) +
  theme(legend.position="none")
```


[REPORT THE RESULTS]
There is a significant effect of diagnosis, b= 0.43, SE = 0.11, t(59) = 3.77, p < .001, and time (VISIT), b=0.12,SE = 0.01, t(292) = 8.69, p < .0001, on the parents' MLU (mean length of utterances).

### Adding new variables (Exercise 4)

Your task now is to figure out how to best describe the children linguistic trajectory. The dataset contains a bunch of additional demographic, cognitive and clinical variables (e.g.verbal and non-verbal IQ). Try them out and identify the statistical models that best describes your data (that is, the children's MLU). Describe how you selected the best model and send the code to run the model to Malte (au540041@post.au.dk).

OPTIONAL: Exercise 5) Comment on how the three linguistic variables measure linguistic performance (the so-called "construct validity" of the measures). Do they express the same variance?


```{r ex4, include = FALSE}
model3 <- lmer(CHI_MLU ~ Diagnosis + Gender + MOT_MLU + poly(VISIT,3) + verbalIQ1 +(1+VISIT|SUBJ) + (1|Ethnicity), data = Data) # 
summary(model3)
#linguistic variables
model4 <- lmer(CHI_MLU ~ types_CHI + tokens_CHI + (1+VISIT|SUBJ), data = Data, REML = FALSE)
summary(model4)
cor(Data$types_CHI, Data$tokens_CHI) #kids using more words use also more types of words - reasonable (types makes more sense to include here, I think.)

#there was sign. effect of verbalIQ, types of words, visit, although not gender, but seems that it should still be a predictor? but not ethnicity,ADOS, non-verbalIQ...

#FINAL MODEL
model5 <- lmer(CHI_MLU ~ VISIT*Diagnosis + MOT_MLU + types_CHI + verbalIQ1 + (1+VISIT|SUBJ), Data, REML=FALSE)
summary(model5)

#evaluate
anova(model5, null_model) 
library("MuMIn")
r.squaredGLMM(model5) #variance explained by fixed factors

```

[REPORT THE RESULTS]
Final model includes the ineraction between time/visit and diagnosis. Also the mothers' MLU seemed to have a significant effect on the children's MLU. I've included the types of words the child uses as a factor but not the amount of words/tokens for linguistic factors. For cognitive abilities, I've included verbal IQ but not non-verbal. + random effect - individual differences per subject visit/time.