---
title: "Assignment2_Part1_VoiceInSchizophrenia"
author: "Réka Keresztényi"
date: "July 17, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pastecs)
library(purrr)
library(rgl)
library(tseriesChaos)
library(SparseM)
library(crqa)
library(stringr)
library(MuMIn)

setwd("C:/Users/Reka/Documents/EM3")
```

## Assignment 2 - Part 1 - Assessing voice in schizophrenia

Schizophrenia has been associated with "inappropriate" voice, sometimes monotone, sometimes croaky. A few studies indicate that pitch might be an index of schizophrenia. However, an ongoing meta-analysis of the literature (which you will have a go at in the last assignment) indicates that pitch mean and standard deviation are only weak indicators of Diagnosis. Can we do better with our new fancy complex skills?

The corecplotus you are asked to analyse is a set of voice recordings from people with schizophrenia (just after first Diagnosis) and 1-1 matched controls (on gender, age, education). Each participant watched 10 videos of triangles moving across the screen and had to describe them (so you have circa 10 recordings per person). We have already extracted the pitch once every 10 milliseconds and you will have to use this data to assess differences in the voice.

Can you characterize voice in schizophrenia as acoustically different? Report the methods you used to answer this question and the results from the analyses. Add a couple of lines trying to interecplotret the results (make sense of the difference). E.g. People with schizophrenia tend to have high-pitched voice, and present bigger swings in their prosody than controls. Add a couple of lines describing limitations of the data/analyses if any is relevant.


N.B. There are looots of files to be dealt with. Maybe too many for your computer, depending on how you load the files. This is a challenge for you. Some (complementary) possible strategies:
- You can select a subset of files only (and you have to justify your choice).
- You can learn how to use the apply() or map() functions.
- You can coordinate with classmates.

Hint: There is some information in the filenames that you might need.
Hint: It might be a good idea to first create a function which loads and parses one file, and then loop through the list of files to read them all. For example

1. In the course of this assignment you have to first select one datafile and figure out how to:

- Extract "standard" descriptors of pitch: Mean, standard deviation, range
- Extract less "standard" descriptors of pitch you can think of (e.g. median, iqr, mean absoluted deviation, coefficient of variation)

```{r, extract info}
#datasets
demo <- read.delim("Demodata.txt")
test <- read.delim("Pitch/Study1D0S101T2_f0.txt") 
art <- read.csv("Articulation.txt")

#Standard descriptors
stat.desc(test_data$time)
stat.desc(test_data$f0)
mean(test_data$f0) #130.60
sd(test_data$f0) #13.76
#sd(test_data$f0)/mean(test_data$f0)*100 #10.54
range(test_data$f0) #114.71 - 229.55
median(test_data$f0) #128.04
#Get aboslute deviation
mad(test_data$time) #2.11
mad(test_data$f0)   #8.88
mad(test_data$f0, center = mean(test_data$f0), na.rm = FALSE) #10.93
#IQR = interquartile range
IQR(test_data$f0) #12.89

#TIME: Mean = 3.86; Standard deviation = 2.91; Range = 0.2 - 8.74; Median = 2.77; Absolute deviation = 2.11; Coefficient of variation = 0.75

#PITCH FREQUENCY: Mean = 130.6; Standard deviation = 13.76; Range = 114.71 - 229.55; Median = 128.04; Absolute deviation = 8.88; Coefficient of variation = 0.11

#Find parameters: delay, dimensions, radius
optimizeParam(test_data$f0,test_data$f0, par = list(lgM = 50, steps = seq(1,6,1),radiusspan = 100, radiussample = 40, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2, tw = 0, whiteline = FALSE, recpt = FALSE, fnnpercent = 10, typeami = "mindip"), min.rec = 3.5, max.rec = 4.5)
#radius: 5.72
#emddim: 3
#delay: 35

#Use the parameters to do the recurrence analysis
x <- crqa(test_data$f0,test_data$f0,embed = 3, delay = 35, radius = 5.717789, rescale = 0, normalize = 0, mindiagline = 2, minvertline = 2)

#recurrence plot 
RP <- x$RP
RP <- matrix(as.numeric(RP), nrow=ncol(RP))
cols <- c("white", "blue")
image(RP, xlab = "", ylab="", col=cols)

```


2. Second you will have to turn the code into a function and loop through all the files (or even better use apply/sapply/lapply)
- Remember to extract the relevant information from the file names (Participant, Diagnosis, Trial, Study)

```{r, loop/extractinfo}
#Create empty list
Subject <- NULL
Study <-  NULL
Trial <-  NULL
Diagnosis <-  NULL
PR <- NULL
DET <- NULL
NRLINE <- NULL
maxL <- NULL
L <- NULL
ENTR <- NULL
LAM <- NULL
TT <- NULL
mean <- NULL
SD <- NULL
range <- NULL
median <- NULL
IQR <- NULL
mad <- NULL
coef_var <- NULL

#loadallfiles
files <- list.files("Pitch/", pattern = ".*.txt", full.names = TRUE)

# for every loop n will be increased by 1 so that we can distinguish between each file
n = 1

#loop
for (file in files) {
  df = read.delim(file)
  Subject[n] <-  substr(file, 14, 18)
  Study[n] <-  substr(file, 7, 12)
  Trial[n] <- sub("_", "", substr(file, 20, 21))
  Diagnosis[n] <-  substr(file, 14, 14)
  mean[n] <- mean(df$f0)
  SD[n] <- sd(df$f0)
  range[n] <- range(df$f0)
  median[n] <- median(df$f0)
  IQR[n] <- IQR(df$f0)
  mad[n] <- mad(df$f0,center = mean(df$f0))
  coef_var[n] <- sd(df$f0)/mean(df$f0)*100
  
  #set parameterlist and optimize them
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  
  parameters <- try(optimizeParam(df$f0, df$f0, par, min.rec = 3.5, max.rec = 4.5))
  
  
  #this if/else statement assures that we the files that does not fit the parameters are still considered in the final lists 
  if (length(parameters) > 1) {
  
  results1 <- crqa(df$f0, df$f0, delay = parameters$delay, embed = parameters$emddim, radius = parameters$radius, normalize = 0, rescale = 0, mindiagline = 2, minvertline = 2)
  
  PR[n] = results1$RR
  DET[n] <- results1$DET
  NRLINE[n] <- results1$NRLINE
  maxL[n] <- results1$maxL
  L[n] <- results1$L
  ENTR[n] <- results1$ENTR
  LAM[n] <- results1$LAM
  TT[n] <- results1$TT
  
  
  }
  
  else {
  
  PR[n] = NA
  DET[n] <- NA
  NRLINE[n] <- NA
  maxL[n] <- NA
  L[n] <- NA
  ENTR[n] <- NA
  LAM[n] <- NA
  TT[n] <- NA
  
  }
  
#increase n by 1 to index into the next value in the lists
  n = n + 1
  
}

all_data <-  data.frame(Subject, Study, Trial, Diagnosis, PR, DET, NRLINE, maxL, L, ENTR, LAM, TT, mean, SD, range, median)

#write csv file with the dataframe 
write.csv(all_data, file = "schiz_data.csv")

```

3. Make one model per acoustic feature and test whether you can observe significant difference due to Diagnosis. Tip: Which other fixed factors should you control for (that is, include in the model)? Which random ones?
- Bonus points: cross-validate the model and report the betas and standard errors from all rounds to get an idea of how robust the estimates are. 
3a. Is Study a significant predictor in these models? What should you infer from this? Does Study interact with Diagnosis? What should you infer from this?
```{r, function}
#merged with demo
data <- read.csv("schizdata.csv")

#Function extracting coefficients and other things from model summary
model_info = function(mdl,n){
  temp_df = summary(mdl)$coefficients
  temp_df = as.data.frame(temp_df)
  temp_df = tibble::rownames_to_column(temp_df, "coef")
  temp_df$R2m = r.squaredGLMM(mdl)[1]
  temp_df$R2c = r.squaredGLMM(mdl)[2]
  temp_df$n = n
  return(temp_df)
}

#mean
model_mean <-model_info(lmerTest::lmer(mean ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),1)
#range
model_range <- model_info(lmerTest::lmer(range ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),2)
#sd
model_sd <-model_info(lmerTest::lmer(SD ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),3)
#median
model_median <-model_info(lmerTest::lmer(median ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),4)
#iqr
model_iqr <-model_info(lmerTest::lmer(IQR ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),5)
#mad
model_mad <-model_info(lmerTest::lmer(mad ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),6)
#coefvar
model_coefvar <-model_info(lmerTest::lmer(coef_var ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),7)
#min
model_min <-model_info(lmerTest::lmer(min ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),8)
#max
model_max <-model_info(lmerTest::lmer(max ~ Diagnosis + Gender  + (1+Trial|Subject) + (1+Trial|Study), data , REML = FALSE),9)

#ALL info together
all_data <- as.data.frame(rbind(model_mean,model_range,model_median,model_max,model_min,model_mad,model_iqr,model_sd,model_coefvar)) #significant all
```

```{r, +study}
#mean+study
mean_model2 <- model_info(lmerTest::lmer(mean ~ Diagnosis + Study + Gender + (1+Trial|Subject), data, REML = FALSE),1)
#sd
sd_model2 <- model_info(lmerTest::lmer(SD ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),2)
#min
min_model2 <- model_info(lmerTest::lmer(min ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),3)
#max
max_model2 <- model_info(lmerTest::lmer(max ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),4)
#median
median_model2 <- model_info(lmerTest::lmer(median ~ Diagnosis + Study + Gender +(1+Trial|Subject),data, REML = FALSE),5)
#iqr
iqr_model2 <- model_info(lmerTest::lmer(IQR ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),6)
#mad
mad_model2 <- model_info(lmerTest::lmer(mad ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),7)
#coefvar
coef_model2 <- model_info(lmerTest::lmer(coef_var ~ Diagnosis + Study + Gender + (1+Trial|Subject) ,data, REML = FALSE),8)
#range
range_model2 <- model_info(lmerTest::lmer(range ~ Diagnosis + Study + Gender + (1+Trial|Subject),data, REML = FALSE),9)

#test
all_data2 = as.data.frame(rbind(mean_model2,sd_model2,min_model2, max_model2, median_model2,iqr_model2,mad_model2,coef_model2,range_model2))

newdata2 <- subset(all_data2, `Pr(>|t|)`<0.05, select=c(n, `Pr(>|t|)`,coef, Estimate, `t value`))

```


```{r, +interaction}
#mean
mean_model3 <- model_info(lmerTest::lmer(mean ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),1)
#sd
sd_model3 <- model_info(lmerTest::lmer(SD ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),2)
#min
min_model3 <- model_info(lmerTest::lmer(min ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),3) 
#max
max_model3 <- model_info(lmerTest::lmer(max ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),4)
#median
median_model3 <- model_info(lmerTest::lmer(median ~ Diagnosis*Study + Gender +(1+Trial|Subject),data, REML = FALSE),5)
#iqr
iqr_model3 <- model_info(lmerTest::lmer(IQR ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),6)
#mad
mad_model3 <- model_info(lmerTest::lmer(mad ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),7)
#coefvar
coef_model3 <- model_info(lmerTest::lmer(coef_var ~ Diagnosis*Study + Gender + (1+Trial|Subject) ,data, REML = FALSE),8)
#range
range_model3 <- model_info(lmerTest::lmer(range ~ Diagnosis*Study + Gender + (1+Trial|Subject),data, REML = FALSE),9)

all_data3 = as.data.frame(rbind(mean_model3,sd_model3,min_model3, max_model3, median_model3,iqr_model3,mad_model3,coef_model3,range_model3))

newdata3 <- subset(all_data3, `Pr(>|t|)`<0.05, select=c(n, `Pr(>|t|)`,coef, Estimate, `t value`))

```


4. Bonus Question: Compare effect size of Diagnosis across the different measures. Which measure seems most sensitive?
- Tip: to compare across measures you need to put all of them on the same scale, that is, you need to "standardize" them (z-score)

5. Bonus question. In the Clinical Info file you have additional information about the participants. Which additional parameters (e.g. age, gender) should we control for? Report the effects.

6. Write a paragraph reporting methods and results

[Next assignment: can we use these measures to build a tool that diagnoses people from voice only?]

## N.B. Remember to save the acoustic features of voice in a separate file, so to be able to load them next time